\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Journal: Ecosphere. Manuscript ID: ECS23-0342. Manuscript type: Article}}

### 2024-02-12 manuscript/manuscriptRW.Rmd
comment 7: figure 7 is too big, create separate figures for each pannel and change references accordingly in the manuscript

In the observed data, recruitment levels decreased clearly with increasing stand basal area (Figure 7a). This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations. The other models featured distinctly different trends, such as (1) an increase in recruitment levels with increasing total basal area followed by the absence of recruitment at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).

The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature. Neither did the models generally match the observed recruitment patterns across these gradients (Figure 7b,c), nor was there a clear pattern across models. Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C. The other models did not feature a clear pattern.

@@

**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. Y-axis are not equally scaled.

**Figure 8.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold. The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient. The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species. The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.

>>>>

In the observed data, recruitment levels decreased clearly with increasing stand basal area (Figure 7). This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations. The other models featured distinctly different trends, such as (1) an increase in recruitment levels with increasing total basal area followed by the absence of recruitment at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).

The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature. Neither did the models generally match the observed recruitment patterns across these gradients (Figure 8,9), nor was there a clear pattern across models. Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C. The other models did not feature a clear pattern.

@@

**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of total basal area. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.

**Figure 8.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of climatic water balance. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.

**Figure 9.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.

**Figure 10.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold. The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient. The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species. The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.
